---
title: "ames_housing_prices_project"
author: "Yamato Kataoka"
date: "12/31/2018"
output: html_document
---

```{r setup, include=FALSE}
# install packages if necessary
list.of.packages <- c("tidyverse","rpart", "randomForest", "modelr", "caret")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# load in packages we'll use
library(tidyverse) # utility functions
library(rpart) # for regression trees
library(randomForest) # for random forests
library(modelr) # for the MAE of model
library(caret) # for spliting the dataset

# set working directory
knitr::opts_knit$set(root.dir = '/cloud/project/ames_housing_prices_project')
```

For this project, we are looking at the data science question:

> Can you figure out how much a house will sell for?

### The Data

#### Accessing Data

Now, we'll download and get the data directly from the competitions data [here](https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data).

The [Ames Housing dataset](http://jse.amstat.org/v19n3/decock.pdf) was compiled by Dean De Cock for use in data science education. It's an incredible alternative for data scientists looking for a modernized and expanded version of the often cited Boston Housing dataset. 

Here's a brief file description;

* train.csv - the training set

* test.csv - the test set

* data_description.txt - full description of each column, originally prepared by Dean De Cock but lightly edited to match the column names used here

* sample_submission.csv - a benchmark submission from a linear regression on year and month of sale, lot square footage, and number of bedrooms

You can find the data dictionaries for on [https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data](https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data)

#### Loading Data

We'll load the data and run the code below to create an object called `iowa_data`.

```{r iowa-data, warning=FALSE}
# You'd read the data and store data in a tibble
iowa_data <- read_csv("data/raw_data/train.csv") 

# we'll make sure Condition1 is a factor & not a char
iowa_data$Condition1 <- as.factor(iowa_data$Condition1)
```

### Exploratory Analysis of Housing Prices Data

```{r summary-data}
# now we'll summarize the iowa_data dataframe
summary(iowa_data)
```

Each little chunk of output (e.g. "X1", "Suburb", "Address", "Rooms") tell us about a specific column in our dataframe.

If the column is numeric, though, it will list information about the mean, median, 25th and 75th quartiles, minimum and maximum.

The median year in which the houses in this dataset were built is 1973 (`YearBuilt`). In addtion, we'll find that the maximum number or rooms are 14 (`TotRmsAbvGrd`).

### Prediction Data Analysis

#### Data Splitting

We'll take our dataset and split it into a `training` set and a `tuning` set. 

```{r splitting-data}
## get Index for training set
set.seed(123)
trainIndex <- createDataPartition(iowa_data$Id, p = .7, 
                                  list = FALSE, 
                                  times = 1)

## split into training and tuning set                                  
iowa_train <- iowa_data %>% slice(trainIndex)
iowa_tune <- iowa_data %>% slice(-trainIndex)
```

#### Decision Tree analysis

We can then fit a new model using our training data and test it using our testing data.

We'll be predicting the `SalePrice` variable. And we start with a narrower set of numeric variables and fit a model that can predict your target variable using the following predictors;

* LotArea

* YearBuilt

* Condition1 (how close to the main road the house is)

* FullBath

* BedroomAbvGr

* TotRmsAbvGrd

We're going to use the `rpart()` function from the `rpart package` to build our decision tree using the prediction target (`SalePrice`) and predictors (`set of numeric variables`).

```{r modeling}
# build a model to predict housing prices in Iowa using our training set
fit <- rpart(SalePrice ~ TotRmsAbvGrd + FullBath + LotArea + Condition1 +
             YearBuilt + BedroomAbvGr, data = iowa_train)
# get the mean average error for our model
mae(model = fit, data = iowa_tune)
```

Then, we just started with one called Mean Absolute Error (also called MAE) for summarizing model quality.

The prediction error for each house is:

> $error=actual−predicted$

With the MAE metric, we take the absolute value of each error. This converts each error to a positive number. We then take the average of those absolute errors. This is our measure of model quality.

We can get the MAE for our model using the `mae()` function, from the `modelr package`. The `mae()` function takes in a model and the dataset to test it against.

On average, our predictions are off by about 37534.09 dollars.

Now, we built prediction model, so we can actually look at the tree it has built.

```{r model}
# plot our regression tree 
plot(fit, uniform=TRUE)
# add text labels & make them 60% as big as they are by default
text(fit, cex=.6)
```

We can now use our fitted model to predict the prices of some houses, using the `predict()` function for our tuning set.

```{r predict, collapse = T}
print("Making predictions for the following 5 houses:")
print(head(iowa_tune))

print("The predictions are")
print(predict(fit, head(iowa_tune)))

print("Actual price")
print(head(iowa_tune$SalePrice))
```

So, as you can above, the first house we predicted cost 181,500 dollars and you predicted it would cost 150,232.8 dollars the error is 31,267.2 dollars.

Then now, we are setting the tree depth with the maxdepth argument to control overfitting vs underfitting.

We can use a utility function to help compare MAE scores from different values for maxdepth:

```{r maxdepth-fanction}
# a function to get the maximum average error for a given max depth. You should pass in
# the target as the name of the target column and the predictors as vector where
# each item in the vector is the name of the column
get_mae <- function(maxdepth, target, predictors, training_data, tuning_data){
    
    # turn the predictors & target into a formula to pass to rpart()
    predictors <- paste(predictors, collapse="+")
    formula <- as.formula(paste(target,"~",predictors,sep = ""))
    
    # build our model
    model <- rpart(formula, data = training_data,
                   control = rpart.control(maxdepth = maxdepth))
    # get the mae
    mae <- mae(model, tuning_data)
    return(mae)
}
```

We can use a for-loop to compare the accuracy of models built with different values for maxdepth. In this case, the lowest MAE is actually 5.

```{r useing-fanction}
# target & predictors to feed into our formula
target <- "SalePrice"
predictors <-  c("TotRmsAbvGrd", "FullBath", "LotArea", "Condition1", "YearBuilt", "BedroomAbvGr")

# get the MAE for maxdepths between 1 & 10
for(i in 1:10){
    mae <- get_mae(maxdepth = i, target = target, predictors = predictors,
                  training_data = iowa_train, tuning_data = iowa_tune)
    print(glue::glue("Maxdepth: ",i,"\t MAE: ",mae))
}
```

37,534.09 is the lowest mean average error for this dataset, which is given this dataset and our current stopping condition, 6 is the maximum number of nodes.

#### Random Forests analysis

```{r modeling_random_forests}
# fit a random forest model to our training set
fitRandomForest <- randomForest(SalePrice ~ TotRmsAbvGrd + FullBath + LotArea + Condition1 +
             YearBuilt + BedroomAbvGr, data = iowa_train)
# get the mean average error for our new model, based on our tuning set
mae(model = fitRandomForest, data = iowa_tune)
```

On average, this predictions are off by around 30,894.18 dollars. This is a big improvement over our previous best decision tree.

