---
title: "Data Tidying Project"
author: "Yamato Kataoka"
date: "11/9/2018"
output: html_document
---

```{r setup, include=FALSE}
## install packages if necessary
list.of.packages <- c("httr", "readxl", "dplyr", "tidyr", "stringr", "reshape2", "skimr", "ggplot2")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

## load packages
library(httr)
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(reshape2)
library(skimr)
library(ggplot2)
```

### Get Data

[Data Set 1](https://data.world/retail/department-store-sales): Sales from the Retail Trade and Food Services Report from the US Census. This dataset only covers Department Stores, though the report covers a wide range of retail types. [1992-2016]

[Data Set 2](https://data.world/garyhoov/retail-sales-growth): US Retail Sales by Store Type with Growth Rate [2009-2014]

```{r data, warning=FALSE} 
#1992-2016
#https://data.world/retail/department-store-sales

GET("https://query.data.world/s/gdk7iwtlisq6vkktmybqqr7hjjty5s", write_disk(tf <- tempfile(fileext = ".xls")))
df1 <- read_excel(tf)

#2009-2014
# https://data.world/garyhoov/retail-sales-growth
GET("https://query.data.world/s/py7kinxvyuxjpzwdjs2ti4wdmui6bi", write_disk(tf <- tempfile(fileext = ".xls")))
df2 <- read_excel(tf)
## the the first row and make that the column names of the data frame
colnames(df2) <- df2[1,]
```

### Save Raw Data

I'm saving the both data objects as df_department.rds and as df_retail.rds. The objects moved into 'data/raw_data'.

```{r save-data, warning=FALSE} 
## use saveRDS() to save each object as a .rds file 
saveRDS(df1, file="df_department.rds")
saveRDS(df2, file="df_retail.rds")
```

### Wrangle Data

I'm wrangling the data which is currently in the wide data format, into a long data format which will have three columns: `business`, `year`, and `n`. 

```{r wrangle-df2, warning=FALSE} 
## an example working with df2
## let's wrangle!
df_retail <- df2 %>%
  ## remove the r from the column names of df2
  magrittr::set_colnames(gsub("r","",df2[1,])) %>% 
  ## add a new column called "business"
  mutate(business = gsub("[…]|[.]","",`Kind of business`)) %>%
  ## filter to include Retail sales or Department stores sales
  filter(grepl('Retail sales, total |Department stores', business)) %>%
  ## only look at columns with year information in them
  select(.,c(matches('19|20'),business)) %>%
  ## take year column and collapse them into a single column
  gather(., "year", "n", 1:(ncol(.)-1)) %>%
  ## make sure the count column `n` is numeric
  mutate(n=as.numeric(n)) %>%
  ## filter to only include the businesses we're interested in
  filter(business == "Retail sales, total "| business=="Department stores ")
```

This code chunk has added the new columns: value and business.

```{r wrangle-df1, warning=FALSE} 
## now, your turn!
## work with df1
df_department <- df1 %>% 
  ## split Period column into one column called "month" and one called "year"
  separate(Period, into = c('month','year'), extra = 'drop', remove = FALSE) %>%
  ## add a column `value` which contains the 
  ## information from the `Value (in millions)` 
  mutate(value = `Value (in millions)`) %>%
  ## group the data frame by the `year` column
  group_by(year) %>%
  ## Summarize the data by creating a new column
  ## call this column `n` 
  ## have it contain the sum of the `value` column
  mutate(n = sum(value)) %>% 
  ### create a new column called `business`
  ## set the value of this column to  be "department stores" 
  ## for the entire data set 
  mutate(business = "department stores") %>%
  ## reorder column names to be : business, year, n
  select(business, year, n)
```

Merging the two data frames.

```{r merge-data, warning=FALSE} 
## Now, combine the two data frames
df_total <- df_retail %>%
  left_join(df_department)
```

This code will plot these three data below.

```{r plot, warning=FALSE, fig.width=10} 
## Plot Retail Sales data
ggplot(df_retail, aes(x=year,y=n,colour=business)) +
  geom_point() 

## Plot Department Sales data
ggplot(df_department, aes(x=year,y=n)) +
  geom_point() 

## Plot Combined Data
ggplot(df_total, aes(x=year,y=as.numeric(n), colour=business)) +
  geom_point()
```
